---
title: "Data Exploration"
author: "Tingting Thompson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

## Data Upload
```{r}
articles <- read_csv("data/articles.csv")
customers <- read_csv("data/customers.csv")
transactions <- read_csv("data/transactions_train.csv")
```


```{r}

dim(articles)
dim(customers)
dim(transactions)

```

# Variable Exploration

## club_member_status

```{r}
sum(is.na(customers$club_member_status))
table(customers$club_member_status)
```

```{r}
ggplot(customers, aes(club_member_status)) + geom_bar()
```


## fashion_news_frequency

```{r}
sum(is.na(customers$fashion_news_frequency))
table(customers$fashion_news_frequency)
```

> "None" and "NONE", seem to be the same, with differing naming, will stick with the capitalized version "None"

```{r}
customers <- customers %>% mutate(fashion_news_frequency = ifelse(fashion_news_frequency=="NONE","None",fashion_news_frequency))
```

> Let's make it a ordered value

```{r}
class(customers$fashion_news_frequency) #want fct
```

```{r}
(customers <- customers %>% mutate(fashion_news_frequency=factor(fashion_news_frequency,ordered = T, levels=c("None","Regularly","Monthly"))))
```

```{r}
ggplot(customers, aes(fashion_news_frequency)) + geom_bar()
```

## age

```{r}

sum(is.na(customers$age))
hist(customers$age)
```



## postal_code

```{r}
sum(is.na(customers$postal_code))
length(unique(customers$postal_code))

```
> 352899 unique ZIP-codes

## t_dat

```{r}
class(transactions$t_dat)
head(transactions$t_dat)

```


### creating a smaller data frame to work with...

```{r}
data.frame(table(transactions$t_dat[100000:160000]))
```

```{r}
t_dat.df <- read_csv("data/t_dat_df.csv")
```


```{r}
#t_dat.df <- data.frame(table(transactions$t_dat))
```

```{r}
#head(t_dat.df)
```
```{r}
#names(t_dat.df) <- c("t_dat","Freq")
```


save for later
```{r}
#write.table(t_dat.df,"data/t_dat_df.csv",sep=",",row.names=F)
```


### initial timeseries
```{r}
ggplot(t_dat.df,aes(t_dat,Freq)) + geom_col()
```

Testing format(<date>) functionality
```{r}

format(t_dat.df$t_dat[1],"%y") == "18"

```

### 2018-2020 timerseries
```{r}
ts.18 <- t_dat.df %>% filter(format(t_dat,"%y")=="18") %>% 
  ggplot(aes(x=t_dat,y=Freq)) + geom_col() + 
  xlim(as.Date("2018-01-01"),as.Date("2018-12-31")) + ylim(0,200000) +
  labs(x="",y="",title="2018")
ts.19 <- t_dat.df %>% filter(format(t_dat,"%y")=="19") %>% 
  ggplot(aes(x=t_dat,y=Freq)) + geom_col() + 
  xlim(as.Date("2019-01-01"),as.Date("2019-12-31")) + ylim(0,200000) +
  labs(x="",y="",title="2019")
ts.20 <- t_dat.df %>% filter(format(t_dat,"%y")=="20") %>% 
  ggplot(aes(x=t_dat,y=Freq)) + geom_col() + 
  xlim(as.Date("2020-01-01"),as.Date("2020-12-31")) + ylim(0,200000) +
  labs(x="Date",y="",title="2020")
gridExtra::grid.arrange(ts.18,ts.19,ts.20)
```










## price

```{r}
sum(is.na(transactions$price))
```

```{r}

transactions %>% ggplot(aes(x="",y=price)) + geom_violin()

```
> it is clear there are outlier prices, that are not suitable for appearance in overall trend. I will be removing these using basic IQR outlier definition

```{r}
Q3 <- quantile(transactions$price)["75%"]
Q1 <- quantile(transactions$price)["25%"]
is_outlier <- function(p) (p > Q3 + (1.5*(Q3-Q1))) | (p < Q1 - (1.5*(Q3-Q1)))

```

```{r}
is_outlier(Q3)
is_outlier(Q1)
is_outlier(0)
is_outlier(-1)
```

```{r}
(inter.trans <- transactions %>% filter(is_outlier(price) != T))
```

```{r}
hist(inter.trans$price)
```
> far more reasonable distribution

```{r}
inter.trans %>% ggplot(aes(price)) + geom_density()

```
### subsetting
> let's create a new data set for price only

```{r}
price.df <- inter.trans %>% select(customer_id,price)

```

```{r}
price.df <- customers %>% select(customer_id,club_member_status,fashion_news_frequency,age) %>% merge(price.df,by="customer_id")
```

```{r}
head(price.df)
```

> let's export this

```{r}
write.table(price.df, "data/prices.csv",sep=",",row.names=F)
```

### grouping price

```{r}
(sub.trans <- transactions %>% 
  select(customer_id,price) %>% 
  group_by(customer_id) %>% 
  summarize(avg_price = mean(price)))
```

```{r}
(customers2 <- merge(customers, sub.trans, by="customer_id"))
```

```{r}
write.table(customers2, "data/customers_wprice.csv",sep=",",row.names=F)
```


## product_type_no

```{r}

```


## product_group_name

```{r}

```


## graphical_appearance_no

```{r}

```


## colour_group_code

```{r}

```



```{r}

head(customers)

```

# bubble graphic

1. generate seasons column in transactions
2. split transactions by seasons & save files
3. isolate by product_type_no, product_group_name, graphical_appearance_no, and color_group_code
4. generate, histogram rankings and/or bubble graphic



