---
title: "Pricing Plots"
author: "Tingting Thompson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```


```{r}
price.df <- read_csv("data/prices.csv")
```

```{r}
price.df <- price.df %>% mutate(fashion_news_frequency = ifelse(fashion_news_frequency=="NONE","None",fashion_news_frequency))
```

```{r}
(price.df <- price.df %>% mutate(fashion_news_frequency=factor(fashion_news_frequency,ordered = T, levels=c("Monthly","Regularly","None"))))
```



### density plots by fashion_news_frequency and club_member_status

```{r}
price.df %>% ggplot(aes(price, color=fashion_news_frequency)) + geom_density()

```

```{r}
price.df %>% ggplot(aes(price, color=club_member_status)) + geom_density()

```

### violin

```{r}
price.df %>% ggplot(aes(y=price, x=fashion_news_frequency)) + geom_violin()
```

```{r}
price.df %>% ggplot(aes(y=price, x=club_member_status)) + geom_violin()
```

### histograms

```{r}
price.df %>% ggplot(aes(price, color=fashion_news_frequency)) + geom_histogram()

```

```{r}
price.df %>% ggplot(aes(price, color=club_member_status)) + geom_histogram()

```



### grid histogram

```{r}
price.df %>% ggplot(aes(price)) + geom_histogram() + facet_wrap(~fashion_news_frequency)

```

```{r}
price.df %>% ggplot(aes(price)) + geom_histogram() + facet_wrap(~club_member_status)

```

```{r}
price.df %>% ggplot(aes(price)) + geom_histogram() + facet_grid(fashion_news_frequency~club_member_status)

```


### grid density

```{r}
price.df %>% ggplot(aes(price)) + geom_density() + facet_wrap(~fashion_news_frequency)

```

```{r}
price.df %>% ggplot(aes(price)) + geom_density() + facet_wrap(~club_member_status)

```

```{r}
price.df %>% ggplot(aes(price)) + geom_density() + facet_grid(fashion_news_frequency~club_member_status) +
  labs(x="Scaled Price", title="Distribution of Prices Amount Varying Club Statuses and News Subscriptions",subtitle="active members tend to have greater spending at lower price ends")

```

> hypothesis: as a brand catering to 'fast fashion', their primary consumers are looking to spend small amounts, frequently, to stay up with the newest fashion. The most active subscrptions, monthly, are a bit more evenly distributed as they perhaps spend regardless of the sale events.


### scatter plot
```{r}

price.df %>% ggplot(aes(x=age,y=price)) + geom_point() + facet_grid(fashion_news_frequency~club_member_status)

```

```{r}

price.df %>% ggplot(aes(x=age,y=price,color=fashion_news_frequency)) + geom_point()

```

```{r}

price.df %>% ggplot(aes(x=age,y=price,color=club_member_status)) + geom_point() 

```

Next steps:
1. mark the high activity points
2. analyze outlier rates and distribution amoung groups


# Hypothesis 1:
whether price is related to purchasing outcome in a sample of N = 1
million after controlling for the effect of customer age, location, and fashion inclination
(fashion_news_frequency).

```{r}

```