---
title: "Seasonal Purchasing Time Series"
author: "Tingting Thompson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggthemes)
```

> idea: split into 2019 and 2020; change color by season (Winter, Spring, Summer, Fall)
> Will need to mutate a new variable for season based on t_dat month
> then split into product_type_no, product_group_name, graphical_appearance_no, and colour_group_code
> convert into shiny object with changing graphics, allow up to 4 graphics on one page?


```{r}
t_dat.df <- read_csv("data/t_dat_df.csv")
```
## Create `season` variable
```{r}
format(t_dat.df$t_dat[1],"%m")

find_season <- function(dt){
  if (format(dt,"%m") %in% c("03","04","05")){
    return("spring")
  }
  else if (format(dt,"%m") %in% c("06","07","08")){
    return("summer")
  }
  else if (format(dt,"%m") %in% c("09","10","11")){
    return("fall")
  }
  else if (format(dt,"%m") %in% c("12","01","02")){
    return("winter")
  }
}

t_dat.df$season <- factor(sapply(t_dat.df$t_dat,find_season),levels=c("fall","winter","spring","summer"),ordered=TRUE)

season.colors <- c(fall="#6E260E",winter="#95ced6",spring="#FB6B90",summer="#F08000")

head(t_dat.df)
```


## seasonal timerseries
```{r}

ggplot(t_dat.df,aes(x=t_dat,y=Freq/1000,fill=season)) + geom_col() +
  scale_color_colorblind() + 
  labs(x="Date", y="# Purchases in Thousands",
       title="Number of Purchases Over Time",
       subtitle="consistent sales besides slight increases in summer and fall, likely back-to-school shopping and gifts purchases for winter holidays (Black Friday)") + 
  theme(legend.position="none", plot.title = element_text(hjust = 0.5),plot.subtitle = element_text(hjust = 0.5)) +
  scale_x_continuous("",breaks=c(as.Date("2018-10-15"),
                                 as.Date("2019-1-15"),
                                 as.Date("2019-4-15"),
                                 as.Date("2019-7-15"),
                                 as.Date("2019-10-15"),
                                 as.Date("2020-1-15"),
                                 as.Date("2020-4-15"),
                                 as.Date("2020-7-15"),
                                 as.Date("2020-10-15")),
                     labels=c("Fall 2018",
                              "Winter 2018",
                              "Spring 2019",
                              "Summer 2019",
                              "Fall 2019",
                              "Winter 2019",
                              "Spring 2020",
                              "Summer 2020",
                              "Fall 2020"))

```

> Hypothesis: Summer and Fall are popular for back-to-school shopping and prepared gifts for Winter holiday season. The obvious black friday surge in the Fall as well.
> There doesn't seem to be enough yearly data to establish any seasonal trends (only 2 full cycles Fall to Fall)

```{r}
ggplot(t_dat.df,aes(x=t_dat,y=Freq,color=as.factor(season))) + geom_point() +
  scale_fill_manual(values=season.colors) 

```

```{r}
ggplot(t_dat.df,aes(x=t_dat,y=Freq,color=as.factor(season))) + geom_line() +
  scale_fill_manual(values=season.colors)
```

## seasonal time series: by year

```{r}
ts.18 <- t_dat.df %>% filter(format(t_dat,"%y")=="18") %>% 
  ggplot(aes(x=t_dat,y=Freq,fill=as.factor(season))) + geom_col() + 
  xlim(as.Date("2018-01-01"),as.Date("2018-12-31")) + ylim(0,200000) +
  labs(x="",y="",title="2018")+ theme(legend.position="none") +
  scale_fill_manual(values=season.colors)
ts.19 <- t_dat.df %>% filter(format(t_dat,"%y")=="19") %>% 
  ggplot(aes(x=t_dat,y=Freq,fill=as.factor(season))) + geom_col() + 
  xlim(as.Date("2019-01-01"),as.Date("2019-12-31")) + ylim(0,200000) +
  labs(x="",y="",title="2019")+theme(legend.position="none",axis.text.y = element_blank()) +
  scale_fill_manual(values=season.colors)
ts.20 <- t_dat.df %>% filter(format(t_dat,"%y")=="20") %>% 
  ggplot(aes(x=t_dat,y=Freq,fill=as.factor(season))) + geom_col() + 
  xlim(as.Date("2020-01-01"),as.Date("2020-12-31")) + ylim(0,200000) +
  labs(x="Date",y="",title="2020")+theme(legend.position="none",axis.text.y = element_blank()) +
  scale_fill_manual(values=season.colors)
gridExtra::grid.arrange(ts.18,ts.19,ts.20, nrow=1)
```

```{r}
ts.18 <- t_dat.df %>% filter(format(t_dat,"%y")=="18") %>% 
  ggplot(aes(x=t_dat,y=Freq,color=as.factor(season))) + geom_point() + 
  xlim(as.Date("2018-01-01"),as.Date("2018-12-31")) + ylim(0,200000) +
  labs(x="",y="",title="2018")+ theme(legend.position="none") +
  scale_fill_manual(values=season.colors)
ts.19 <- t_dat.df %>% filter(format(t_dat,"%y")=="19") %>% 
  ggplot(aes(x=t_dat,y=Freq,color=as.factor(season))) + geom_point() + 
  xlim(as.Date("2019-01-01"),as.Date("2019-12-31")) + ylim(0,200000) +
  labs(x="",y="",title="2019")+theme(legend.position="none",axis.text.y = element_blank()) +
  scale_fill_manual(values=season.colors)
ts.20 <- t_dat.df %>% filter(format(t_dat,"%y")=="20") %>% 
  ggplot(aes(x=t_dat,y=Freq,color=as.factor(season))) + geom_point() + 
  xlim(as.Date("2020-01-01"),as.Date("2020-12-31")) + ylim(0,200000) +
  labs(x="Date",y="",title="2020")+theme(legend.position="none",axis.text.y = element_blank()) +
  scale_fill_manual(values=season.colors)
gridExtra::grid.arrange(ts.18,ts.19,ts.20)
```

## Next Steps:
> Now we need to control for these variables: product_type_no, product_group_name, graphical_appearance_no, and colour_group_code.
> Let's create a sub plotting of this
> easiest: swimsuits vs outerwear sales


